---
title: "Introduction to fqardl: Fourier Quantile and Nonlinear ARDL"
author: "Muhammad Alkhalaf"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to fqardl}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

The `fqardl` package provides R implementations of advanced ARDL (Autoregressive 
Distributed Lag) methodologies for cointegration analysis:

1. **FQARDL**: Fourier Quantile ARDL - combines quantile regression with Fourier 
   approximation for structural breaks
2. **FNARDL**: Fourier Nonlinear ARDL - captures asymmetric effects with Fourier 
   terms for smooth breaks

## Installation

```{r eval=FALSE}
# Install from CRAN
install.packages("fqardl")

# Or from GitHub
# devtools::install_github("muhammedalkhalaf/fqardl")
```

## Quick Start

```{r}
library(fqardl)
```

### Generate Example Data

```{r}
set.seed(123)
n <- 150

# Create cointegrated series with structural break
t <- 1:n
x1 <- cumsum(rnorm(n)) + 0.3 * sin(2 * pi * t / n)
x2 <- cumsum(rnorm(n))

y <- numeric(n)
y[1] <- 0
for (i in 2:n) {
  # Relationship changes at t = 75
  if (i < 75) {
    y[i] <- 0.3 * y[i-1] + 0.5 * x1[i] + 0.3 * x2[i] + rnorm(1, sd = 0.5)
  } else {
    y[i] <- 0.3 * y[i-1] + 0.8 * x1[i] + 0.2 * x2[i] + rnorm(1, sd = 0.5)
  }
}

data <- data.frame(y = y, x1 = x1, x2 = x2)
```

### FQARDL: Quantile Analysis

```{r eval=FALSE}
# Estimate FQARDL across quantiles
result <- fqardl(
  formula = y ~ x1 + x2,
  data = data,
  tau = c(0.25, 0.50, 0.75),
  max_k = 3,
  criterion = "BIC"
)

summary(result)
```

### FNARDL: Asymmetric Effects

```{r eval=FALSE}
# Create data with asymmetric relationship
oil <- cumsum(rnorm(n, sd = 2))
gdp <- numeric(n)
gdp[1] <- 100

for (i in 2:n) {
  d_oil <- oil[i] - oil[i-1]
  if (d_oil > 0) {
    gdp[i] <- 0.95 * gdp[i-1] - 0.3 * d_oil + rnorm(1, sd = 0.5)
  } else {
    gdp[i] <- 0.95 * gdp[i-1] - 0.7 * d_oil + rnorm(1, sd = 0.5)
  }
}

asym_data <- data.frame(gdp = gdp, oil = oil)

# Estimate FNARDL
result_asym <- fnardl(
  formula = gdp ~ oil,
  data = asym_data,
  decompose = "oil",
  max_k = 2
)

summary(result_asym)
```

## Methodology

### Fourier Approximation

The package uses Fourier trigonometric terms to capture smooth structural breaks:

$$sin(2\pi k t / T), \quad cos(2\pi k t / T)$$

where $k$ is the Fourier frequency and $T$ is the sample size. This approach, 
following Enders & Lee (2012), allows detecting breaks without specifying 
their timing.

### Quantile ARDL

FQARDL estimates the model at different quantiles $\tau \in (0,1)$:

$$Q_\tau(y_t | X_t) = \alpha(\tau) + \phi(\tau) y_{t-1} + \beta(\tau)' X_{t-1} + \ldots$$

This reveals how relationships vary across the conditional distribution.

### Nonlinear ARDL

FNARDL decomposes regressors into positive and negative partial sums:

$$x_t^+ = \sum_{j=1}^{t} \max(\Delta x_j, 0), \quad x_t^- = \sum_{j=1}^{t} \min(\Delta x_j, 0)$$

This captures asymmetric effects following Shin, Yu & Greenwood-Nimmo (2014).

## References

- Cho, J. S., Kim, T., & Shin, Y. (2015). Quantile cointegration in the 
  autoregressive distributed-lag modeling framework. *Journal of Econometrics*.

- Enders, W., & Lee, J. (2012). A unit root test using a Fourier series to 
  approximate smooth breaks. *Oxford Bulletin of Economics and Statistics*.

- Pesaran, M. H., Shin, Y., & Smith, R. J. (2001). Bounds testing approaches 
  to the analysis of level relationships. *Journal of Applied Econometrics*.

- Shin, Y., Yu, B., & Greenwood-Nimmo, M. (2014). Modelling asymmetric 
  cointegration and dynamic multipliers in a nonlinear ARDL framework. 
  In *Festschrift in Honor of Peter Schmidt*. Springer.
